---
title: "TidyVerse CREATE"
author: "Arushi Arora"
date: "10/23/2020"
output: openintro::lab_report
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introdction
The following example from Five Thirty Eight https://fivethirtyeight.com/features/how-americans-like-their-steak/

The core tidyverse package includes "readr" and "dplyr"
"readr" provides a fast and friendly way to read rectangular data
"dplyr" provides a grammar of data manipulation, providing a consistent set of verbs that solve the most common data manipulation 

#### Importing libraries
```{r warning=FALSE}
if(!require('tidyverse'))install.packages('tidyverse')
library (tidyverse)
library(dplyr, warn.conflicts = FALSE) # for cleaning the data
options(dplyr.summarise.inform = FALSE)
```


```{r warning=FALSE}
urlfile="https://raw.githubusercontent.com/fivethirtyeight/data/master/steak-survey/steak-risk-survey.csv"

steakdata<- readr::read_csv(url(urlfile))

head(steakdata)
```

## Rename columns

```{r warning=FALSE}
steakdata1 = dplyr::rename(steakdata, 
"lottery" = "Consider the following hypothetical situations: <br>In Lottery A, you have a 50% chance of success, with a payout of $100. <br>In Lottery B, you have a 90% chance of success, with a payout of $20. <br><br>Assuming you have $10 to bet, would you play Lottery A or Lottery B?", 
"smoke_cigs" = "Do you ever smoke cigarettes?" ,
"drink_alcohol" = "Do you ever drink alcohol?", 
"gamble" = "Do you ever gamble?",
"skydiving" = "Have you ever been skydiving?",
"overspeeding" = "Do you ever drive above the speed limit?",
"cheat_patner" = "Have you ever cheated on your significant other?",
"eat_steak" = "Do you eat steak?",
"steak_prep" = "How do you like your steak prepared?",
"hh_income" = "Household Income",
"location" = "Location (Census Region)")
head(steakdata1)
```

## Remove first row

```{r warning=FALSE}
steakdata2 <- steakdata1[-c(1), ]

head(steakdata2)
```

## Mutate Char variables to Factor

```{r warning=FALSE}
steakdata3 <- steakdata2 %>% as_tibble() 

steakdata4 <- steakdata3 %>%
  mutate(lottery = as.factor(lottery)) %>%
  mutate(smoke_cigs = as.factor(smoke_cigs)) %>%
  mutate(drink_alcohol = as.factor(drink_alcohol)) %>%
  mutate(gamble = as.factor(gamble)) %>%
  mutate(skydiving = as.factor(skydiving)) %>%
  mutate(overspeeding = as.factor(overspeeding)) %>%
  mutate(cheat_patner = as.factor(cheat_patner)) %>%
  mutate(eat_steak = as.factor(eat_steak)) %>%
  mutate(steak_prep = as.factor(steak_prep)) %>%
  mutate(Gender = as.factor(Gender)) %>%
  mutate(Age = as.factor(Age)) %>%
  mutate(hh_income = as.factor(hh_income)) %>%
  mutate(Education = as.factor(Education)) %>%
  mutate(location = as.factor(location))
```


# Tidyverse Extend

Karim here to add some further analysis, Arushi has done a great job analyzing How Americans Like Their Steak. 

```{r}
#Lets remove the N/A from the following columns
steakdata4 <- subset(steakdata4, Age != "N/A")
steakdata4 <- subset(steakdata4, Education != "N/A")
steakdata4 <- subset(steakdata4, hh_income != "N/A")

head(steakdata4)
```


## Subset distinct observations from the data

```{r}
#remove respondentID column
steakdata5 <- steakdata4[2:15]
steakdata6 <- dplyr::distinct(steakdata5)

head(steakdata6)
```

## Extract a single colum from data
```{r}
#extract drink_alcohol column by index
steakdata7 <- dplyr::pull(steakdata6, 3)

#extract drink_alcohol column by negative index
steakdata7 <- dplyr::pull(steakdata6, -12)

#extract drink_alcohol column by name
steakdata7 <- dplyr::pull(steakdata6, drink_alcohol)

head(steakdata7)
```

## Change the column order

```{r}
#move Gender column to second column using .before
steakdata8 <- dplyr::relocate(steakdata4, Gender, .before = lottery)

#move Gender column to second column using .after
steakdata8 <- dplyr::relocate(steakdata4, Gender, .after = RespondentID)

head(steakdata8)
=======
Lets check how Age, Education, and steak_prep by building a animated bubble chart using the 'gganimate' library to extend ggplot2's functionality.

```{r, message = FALSE}
# package to extend ggplot2 functionality
library(gganimate)
# package that will display the animated output when you knit to HTML
library(gifski)
```


```{r warning=FALSE}
# Setup basic ggplot barchart
ggplot(steakdata4, aes(x= Age, y=Education, fill=Education)) + 
  geom_bar(stat='identity') +
  theme(legend.position="bottom", legend.box="vertical", legend.margin=margin()) +
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  guides(fill=guide_legend(nrow=2,byrow=TRUE)) +
  coord_flip() +
  
# steak_prep as the variable it transitions on
  transition_states(steak_prep) +
  ease_aes('sine-in-out') + 
  labs(title = 'How Americans like their steak: {closest_state}', x = 'Age', y = 'Level of Education')
```
