---
title: "Tidyverse CREATE Assignment"
author: "Peter (CREATE), and Cameron Smith (EXTEND)"
date: "10/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Tidyverse CREATE section

In this assignment, you’ll practice collaborating around a code project with GitHub.  You could consider our collective work as building out a book of examples on how to use TidyVerse functions.

GitHub repository:  https://github.com/acatlin/FALL2020TIDYVERSE

https://data.fivethirtyeight.com/ datasets.

https://www.kaggle.com/datasets datasets. 

Your task here is to <b>Create an Example.</b>  Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset. (25 points)

## Loading of required libraries

```{r}
library(tidyverse)
```

## Dataset

https://www.kaggle.com/ronitf/heart-disease-uci

## File on Github

https://raw.githubusercontent.com/petferns/607-week9/main/heart.csv

## Capability1 readr::read_csv

First lets try to read the CSV file from GitHub using the read_csv function

```{r}
heart <- read_csv("https://raw.githubusercontent.com/petferns/607-week9/main/heart.csv")
head(heart)
```

read_csv function from tidyverse library is faster than the default read.csv from R. Let us verify the same

```{r}
system.time(d<-read.csv("https://raw.githubusercontent.com/petferns/607-week9/main/heart.csv"))
```


```{r message=FALSE}
system.time(d<-read_csv("https://raw.githubusercontent.com/petferns/607-week9/main/heart.csv"))

```

## Capability 2 dplyr::filter

As the name suggests we can filter the rows from the dataframe using the filter function.
We apply the filter to heart dataframe based on age column

```{r}
filter(heart,age > 45)
```

## Capability 3 dplyr::select

Just as we filtered the rows of a dataframe in the above example, we can use select function to filter columns of the dataframe. If you want only specific columns rather than the whole set of columns we can use select function

```{r}
select(heart,c("age","sex","chol"))
```

# Tidyverse EXTEND Section

Extend Section Author: Cameron Smith
Date Submitted: 7 November 2020
Quick Summary: Added mutate and transmute examples and cleaned updated the overall RMD structure a bit.

## Introduction

For the Data 607 Tidyverse Extend assignment I am extending the original vignette to demonstrate more functionality.  As demonstrated above in the original section with the read_csv, select, and filter functions from the Tidyverse, there are quite a few functions (arguably all) in the Tidyverse that largely mirror base R functions, but with slightly different syntax and intent.  

Additional good examples of this are the 'mutate' and 'transmutate' functions which I will demonstrate below.

## Dataset

For this section I have used the Netflix Titles dataset from Kaggle, link and info below.  I have copied it into and loaded it from my personal Github account for reproducibility.

https://www.kaggle.com/shivamb/netflix-shows
Version 3, Accessed 7 November 2020.

## Load the Data

```{r}

# Loading data from Github into new data frame
df_extend <- read.csv("https://raw.githubusercontent.com/cwestsmith/cuny-msds/master/datasets/netflix_titles.csv")

# Quick look at the new data frame to make sure it loaded ok
str(df_extend)

# Trim the data frame down a bit
df_extend <- df_extend %>% select(type, title, release_year)

# New look at the data frame to verify changes
str(df_extend)
```

## Capability 4 dplyr::mutate

As seen above, the data frame currently only has 3 columns: type, title, and release_year.  What if we wanted to add a new column based on the age of the movie (relative to the current year)?  Mutate is a good tool for this.  It adds new columns using a straightforward, simple syntax and preserves the old ones.  Example below.

```{r}
# Create variable containing the current year
current_year <- as.numeric(format(Sys.time(), "%Y"))

# Add new column and view the results for the first 5 movies
# We can see that all original columns are still there, including the new one
df_extend %>% 
  mutate(movie_age = current_year - release_year)
```

## Capability 5: dplyr::transmute

Transmute is very similar to mutate in that it is used to create new columns.  However, transmute does not preserve the original columns.  Example below.

```{r}

# Add new column with a shortened title and view the results
# In this example we are using another Tidyverse function from the stringr package to shorten a string.

# Unlike with mutate, we can see below that the original columns have been discarded and only the new column remains.  Essentially it acts like the select() function, only selecting the new column created.
df_extend %>% 
  transmute(short_title = str_trunc(title, width = 20))

```

## Conclusion for Tidyverse EXTEND section

In addition to the functions highlighted in the CREATE section of this document, mutate() and transmute() are very useful.  They are very similar in what they do with one key difference - mutate preserves the original columns and transmute does not.
