---
title: "TidyVerse EXTEND Assignment"
author: "Jered Ataky"
date: "2020-11-6"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Highlight

In this "Tidyverse EXTEND" recipe,
we are going to extend the work created by "Zhouxin Shi".

There will be two parts: "The original recipe: tidyverse CREATE" which is the vignette
created by Zhouxin, and "Tidyverse EXTEND" which is our
additional work (code) to the original recipe.


## 1. Original recipe: tidyverse CREATE


### Tidyverse

In this assignment, you’ll practice collaborating around a code project with GitHub.  You could consider our collective work as building out a book of examples on how to use TidyVerse functions.

GitHub repository:  https://github.com/acatlin/FALL2020TIDYVERSE

FiveThirtyEight.com datasets.

Kaggle datasets. 

Your task here is to Create an Example.  Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset. (25 points)


### Read data using readr::read_csv

```{r }
library(tidyverse)

pums <- read_csv("https://raw.githubusercontent.com/szx868/data607/master/Tidyverse/2019PUMS_PERSON_DATA_NY.csv")


```
### Using dplyr::select to have column you needed

```{r}
select(pums,c("Age","SEX","Total_Personal_Earnings","Total_Personal_Income"))

```

### Using dplyr::filter to age

```{r}
filter(pums,Age > 36)

```

#### Select with blacklist vs conventional whitelist
Lets extend the 'Select' function we just covered. We can use this function to select a subset of columns but also excluding explicit columns. Sometimes this leads to less syntax, especially when you want to keep a majority of the columns.

On top of that we will also include the `head` function so that we can view a sample of our data, without running the entire dataframe into memory.

```{r}
pums %>%
  select(-Worker_Class, -School_Attending) %>%
    head()
```

#### Renaming Columns with Select
Select can also be used to rename columns. However, that drops all the columns that are not mentioned. 


```{r}
pums %>%
  select(gender = SEX,
         salary = Total_Personal_Earnings,
         other_income = Total_Personal_Income
         ) %>%
  head()

```

##### Select a subset and rename
But what if you only want to rename some of the columns and also maintain a subset of the data? In that case, you can use the rename function.

```{r}
pums %>%
  # selects subset of original dataframe
  select(-Worker_Class, -School_Attending) %>%
  # renames columns with new header
    rename(gender = SEX,
           salary = Total_Personal_Earnings,
           other_income = Total_Personal_Income
           ) %>%
  # displays only first 5 rows
      head(5)


```
 
=======
## 2. Tidyverse EXTEND


### Introduction

As this vignette is about dplyr, we are going to extend this work by adding 
different other functions of the same package.
The first part explored the functions select() and filter(), here we are going 
to add three other main verbs of dplyr: arrange(), mutate(), and summarize().

Throughout this second part of the vignette, we make use of the subset of pums
data set created using select() function as in part 1 above.


```{r}

df <- pums %>%
  select("Age","SEX","Total_Personal_Earnings","Total_Personal_Income")

```

### arrange(): Reordering the cases

Arrange function reorders the cases in the order that you want.
Let say you want to reorder pums data frame descending order, you might write: 

Arrange function lets reordering the cases in the order that you want.
Let say that you want to reorder the head of previous selected variables data frame (data2)
in descending order of students math score, you might write:

```{r}

# Reorder in descending order of math score

df %>%
  arrange(desc(Age))
```


### mutate(): creating new variables that are functions of existing variables

The mutate function creates new variables that are functions of the existing variables. 
Let say you want to create "Total_Other_Sources_of_Income" variable which is the difference
between "Total_Personal_Income" and "Total_Personal_Earnings", you might write:

```{r}
df %>%
  mutate(Total_Other_Sources_of_Income 
         = Total_Personal_Income - Total_Personal_Earnings)
```


Let say you are interesting in keeping only the new variable created from the existing variables,
meaning keeping only "Total_Other_Sources_of_Income" variable but neither
"Total_Personal_Income" nor "Total_Personal_Earnings",
you might use another function called transmuse():

```{r}
df1 <- df %>%
  transmute(Age, SEX, 
            Total_Other_Sources_of_Income = Total_Personal_Income - Total_Personal_Earnings )
df1
```

### summarize(): summarizing multiple values to a single value

summarize() will be used with group_by function group_by which helps grouping the data set by a variable. 
Let say you are interested in the summary of the average total personal income by sex, you might write:

```{r}

# Note that we remove missing values in the calculation to calculate the average

df %>%
  group_by(SEX) %>%
  summarize(average_total_income = 
              sum(Total_Personal_Income, na.rm = TRUE) / n())
```

